The Influence of Radial Drift and Gas Accretion on the C/O Ratio in Protoplanetary Disks

### 2016-04-13 intro.tex
upd

\chapter{Introduction}
\label{introduction}

Within the last two decades, more than one thousand extrasolar planets (exoplanets)
have been discovered [1]. Their diversity in terms of mass, radius, location and composition
[2] provides an exciting field of research, with the eventual goal of finding planets that are
similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore
and understand how planets obtain their compositions. Observations of Earth-like planets
that can provide useful insight about their composition are challenging — the solid interior
structure of terrestrial planets cannot be detected, and their gaseous envelopes are small by
comparison (both in mass and radius), which makes it difficult to obtain atmospheric spectra
and find out what chemical compounds they are made of. We therefore turn to giant planets,
which have provided a rich and intriguing research area for decades. Gas giants contain most
of their mass in their atmosphere, hence their chemical composition is determined by that
of their envelopes. The last few years have seen a substantial increase in the number of
giant planets with observed atmospheric spectra (e.g., [3], [4]), which has enhanced our
understanding of these planets’ chemical structure, and has provided us with quantitative
information about the abundances of various compounds in their envelopes besides hydrogen
and helium. Finally, gas giants shape the architecture of planetary systems and affect the
delivery of volatile compunds to terrestrial planets, which has direct consequences for the
habitability of worlds similar to our own. Thus testing theories of planet formation against
gas giant compositions will help constrain planet formation theories more generally.
Both terrestrial and giant planets are born in protoplanetary disks, which implies that
their compositions are determined by and tightly linked to the structure and
composition of the disk. The chemical and dynamical evolution of disks, as well the
formation of giant planets have both been previously investigated in isolation. However, the
coupled chemo-dynamical disk evolution, planet compositions, and most importantly the
disk-planet connection have not yet been considered in detail. In this thesis, I uncover some of the answers to this issue from two standpoints: (1) by looking at the role of disk location in setting the conditions for the formation of wide-separation gas giants, and (2) by investigating how the structure and chemical composition of the protoplanetary disk at different radii affects the composition of nascent giant planets.

@@

>>>>

\documentclass[apj, numberedappendix]{emulateapj}

\usepackage{graphicx} 
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{subeqnarray}
\usepackage{epstopdf}
%\usepackage{natbib}
%\bibliographystyle{apj}

\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}

\newcommand{\ie}{i.e.\ }
\newcommand{\eg}{e.g.\ }
\newcommand{\p}{\partial}
\newcommand{\xv}{\vc{x}}
\newcommand{\kv}{\vc{k}}
\newcommand{\brak}[1]{\langle #1\rangle}

\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
\newcommand{\cm}{\; {\rm cm}}
\newcommand{\mm}{\; {\rm mm}}
%\newcommand{\ps}{\; {\rm s^{-1}}}
\newcommand{\km}{\; {\rm km}}

\newcommand{\AU}{\; {\rm AU}}
\newcommand{\yr}{\; {\rm yr}}
\def\K{\; {\rm K}}

\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
\newcommand{\nab}{\vc{\nabla}}
\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}

\newcommand{\eq}[1]{equation\,(\ref{#1})}
\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
\newcommand{\Eqsss}[3]{Equations (\ref{#1}), (\ref{#2}) and~(\ref{#3})}
\newcommand{\App}[1]{Appendix~\ref{#1}}
\newcommand{\Sec}[1]{Sect.~\ref{#1}}
\newcommand{\Chap}[1]{Chapter~\ref{#1}}
\newcommand{\Fig}[1]{Figure~\ref{#1}}
\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
\newcommand{\Tab}[1]{Table \ref{#1}}

\newenvironment{packed_item}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

\newcommand{\delad}{\nabla_{\rm ad}}
\newcommand{\delrad}{\nabla_{\rm rad}}
\newcommand{\Rg}{\mathcal{R}}
\newcommand{\RB}{R_{\rm B}}
\newcommand{\RH}{R_{\rm H}}
\newcommand{\co}{_{\rm c}}
\newcommand{\pla}{_{\rm p}}
\newcommand{\di}{_{\rm d}}
\newcommand{\cb}{_{\rm RCB}}
\newcommand{\mc}{m_{\rm c \oplus}}
\newcommand{\mcn}[1] { m_{ \rm c #1} }
\newcommand{\mpn}[1] { m_{ \rm p #1} }
\newcommand{\MC}{M_{\rm crit}}
\newcommand{\au}{a_\oplus}
\newcommand{\aun}[1]{ a_{#1} }
\def\crit{_{\rm{c, crit}}}

\begin{document}

\section{Introduction}

Two big questions. In the context of core accretion:
- where in the disk can you form Gas Giants?
- what compositions will the formed Gas Giants obtained?

we can learn about this through a combination of studying planet formation in protoplanetary disks, and the end-results of planet formation, i.e. exo-planets and the architecture and composition of our own Solar System

1. Protoplanetary disks
1a. Density and temperature structures
1b. Chemical processes and structures

2. Observational constraints on Gas Giant formation and compositions

3. Core accretion formation of planets and its challenges

4. Interactions between disk physics and chemistry, and core accretion

5. Connection between disk compositions and the compositions of nascent planets

In recent years, the field of giant planet formation has aimed to answer two fundamental questions:

\begin{enumerate}
\item Where in the protoplanetary disk can gas giants form?
\item What compositions will the formed giant planets have obtained?
\end{enumerate}

We can start uncovering the answers to these questions through a combination of studying planet formation in protoplanetary disks, the end-results of planet formation (i.e., exoplanets), and through comparisons with the architecture and composition of our own Solar System. Within the last two decades, more than one thousand extrasolar planets (exoplanets)
have been discovered \citet{batahla14}. Their diversity in terms of mass, radius, location and composition
\citet{lissauer14} provides an exciting field of research, with the eventual goal of finding planets that are
similar to our own Earth and may sustain life. For this purpose, it is thus crucial to explore
and understand how planets obtain their compositions. Observations of Earth-like planets
that can provide useful insight about their composition are challenging — the solid interior
structure of terrestrial planets cannot be detected, and their gaseous envelopes are small by
comparison (both in mass and radius), which makes it difficult to obtain atmospheric spectra
and find out what chemical compounds they are made of. We therefore turn to giant planets,
which have provided a rich and intriguing research area for decades. Gas giants contain most
of their mass in their atmosphere, hence their chemical composition is determined by that
of their envelopes. The last few years have seen a substantial increase in the number of
giant planets with observed atmospheric spectra (e.g., \citealt{debes13}, \citealt{kreidberg15}), which has enhanced our
understanding of these planets’ chemical structure, and has provided us with quantitative
information about the abundances of various compounds in their envelopes besides hydrogen
and helium. Finally, gas giants shape the architecture of planetary systems and affect the
delivery of volatile compunds to terrestrial planets, which has direct consequences for the
habitability of worlds similar to our own. Thus testing theories of planet formation against
gas giant compositions will help constrain planet formation theories more generally.

Both terrestrial and giant planets are born in protoplanetary disks, which implies that
their compositions are determined by and tightly linked to the structure and
composition of the disk. The chemical and dynamical evolution of disks, as well the
formation of giant planets have both been previously investigated in isolation. However, the
coupled chemo-dynamical disk evolution, planet compositions, and most importantly the
disk-planet connection have not yet been considered in detail. In this thesis, I uncover some of the answers to this issue from two standpoints: (1) by looking at the role of disk location in setting the conditions for the formation of wide-separation gas giants, and (2) by investigating how the structure and chemical composition of the protoplanetary disk at different radii affects the composition of nascent giant planets.

\subsection{Protoplanetary Disks}

Chemical and dynamical processes in a protoplanetary disk affect the disk structure
and composition, and thus the composition of nascent planets. Figure \ref{fig:disk_henning} shows various dynamical and chemical processes that occur in disks, many of which are not yet fully understood. Moreover, the timescales for various disk chemical and dynamical processes may be comparable at
least in some parts of the disk (see Figure \ref{fig:timescales_semenov}).

@@

\end{document}
